### 2017-10-15 mcmc.tex
symbols

Given this pdf $p(\pars)$, we can define \emph{expectation values}
$E_{p(\pars)}[\pars]$
for $\pars$ or for any quantity that can be expressed as a function $q(\pars)$
of $\pars$:
\begin{eqnarray}
E_{p(\pars)}[\pars] &\equiv& \int \pars\,p(\pars)\,\dd \pars
\\
E_{p(\pars)}[q(\pars)] &\equiv& \int q(\pars)\,p(\pars)\,\dd \pars
\quad ,
\end{eqnarray}
where again the integrals are implicitly definite integrals over the
entire domain
of $\pars$ (all the parts of $\pars$ space in which $p(\pars)$ is finite) and the
integrals are multi-dimensional if $\pars$ is multi-dimensional.
These expectation values are the mean values of $\pars$ and $q(\pars)$ under
the pdf.  A good sampling---and really this is the definition of a good
sampling---makes the sampling approximation to these integrals
accurate.
With a good sampling $\setof{\pars_k}_{k=1}^K$ the integrals get replaced
with sums over samples $\pars_k$:
\begin{eqnarray}
E_{p(\pars)}[\pars] &\approx& \frac{1}{K}\,\sum_{k=1}^K \pars_k
\\
E_{p(\pars)}[q(\pars)] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\quad .
\end{eqnarray}
That is, a sampling is good when any expectation value of interest is
accurately computed via the sampling approximation.
The word ``accurately'' here translates into some kinds of theorems
about limiting behavior; the general idea is that the sampling
approximation becomes exact as $K$ goes to infinity.
The size $K$ of the sampling you need \emph{in practice} will depend
on the expectations you want to compute, and the accuracies you need.

@@

When the sampling $\setof{\pars_k}_{k=1}^K$ is of one of these badly
normalized functions $f(\pars)$---as it usually will be---the
sampling-approximation expectation values are the expectation values
under the properly-normalized corresponding pdf, even though you might
\emph{never learn that normalization}.
When we run \MCMC\ sampling on $f(\pars)$, a function that differs from a
pdf $p(\pars)$ by some unknown normalization constant, then the sampling
permits computation of the following kinds of quantities:
\begin{eqnarray}
E_{p(\pars)}[q(\pars)] &\equiv& \frac{\int q(\pars)\,f(\pars)\,\dd \pars}{\int f(\pars)\,\dd \pars}
\\
E_{p(\pars)}[q(\pars)] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\label{eq:expectationvalue}
\quad .
\end{eqnarray}
That is, the sampling can be constructed (as we will show below in \sectionname~\ref{sec:MH}) from
evaluations of $f(\pars)$ directly, and it permits you to compute expectation values
without ever requiring you to integrate either the numerator integral
or the denominator integral, both of which are generally
intractable.\note{As we will
  see, the intractability comes from various directions, but one is
  that the dimension of $\pars$ gets large in most realistic situations.
  Another is that the support of $p(\pars)$ tends to be, in normal
  inference situations, \emph{far smaller} than the full domain of
  $\pars$.
  That is, the pdf is at or very close to zero over all but some tiny
  and very hard-to-find part of the space.}

@@

That is, if the expectation value $E[q]$ is computed by the sampling
approximation given in \equationname~(\ref{eq:expectationvalue}),
there is an uncertainty on that expectation given by
\begin{eqnarray}
  \Var[E_{p(\pars)}[q(\pars)]] &=& \frac{\tau_q}{K}\,\Var_{p(\pars)}[q(\pars)]
  \\
  &\approx& \frac{\tau_q}{K^2}\,\sum_{k=1}^K \left(q(\pars_k) -
    E_{p(\pars)}[q(\pars)]\right)^2
  \quad,
\label{eq:tauvar}
\end{eqnarray}
where we are being a bit loose with terminology because this is the
variance not on $E_{p(\pars)}[q(\pars)]$ but on our sampling-based
\emph{estimator} for
$E_{p(\pars)}[q(\pars)]$, and the $E_{p(\pars)}[q(\pars)]$ inside the sum is
that estimator too.
It is slightly off-topic here, but related to this, if you have a long
chain of length $K$ and with autocorrelation time $\tau$ and you thin
the chain down to every $\tau$th sample (instead of using all $K$
samples), you will get just as good sampling-based estimates as if you
had used all $K$ samples.

It is also worth noting that there is not just one autocorrelation function or
autocorrelation time for an \MCMC\ chain.
Instead, there is an autocorrelation time $\tau_f$ for every function
$f(\pars)$ and these will, in general, be different.
Therefore, the autocorrelation time used in \equationname~(\ref{eq:tauvar}) is
the time computed on the chain of $q(\pars)$ values instead of the
autocorrelation time for $\pars$.

@@

For example,
  imagine that by \MCMC\ you have generated
  $K$ samples $\pars_k$ of a parameter vector $\pars$.
Now you have a scalar function $q(\pars)$ which takes the parameter vector
  and returns a scalar value%
  \note{For the median-of-sampling value, $q(\pars)$ needs to return a scalar,
  but technically, for the mean-of-sampling, $q(\pars)$ can be a vector or something high-dimensional.}
  which could be as simple as a single component of the parameter vector
  (one parameter from the list),
  or something more complex.
The mean-of-sampling value $\mean{q}$ is just
\begin{eqnarray}
\mean{q} &\leftarrow& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\quad ,
\end{eqnarray}
  and the median-of-sampling value is just the $[K/2]$th value of $q(\pars_k)$
  when the $q(\pars_k)$ have been ordered (sorted).
These are both produced by integrals;
  the first is the value returned by the expectation integral estimate,
  the second is the value past which half of the integrated pdf lies.

What are your best options for the ``uncertainty'' or ``limits'' on this parameter?
The only simple integral-based options are either variances or else posterior quantiles.
We usually use quantiles.
That is, the ``one-sigma'' error bar can be taken to be the half-size of the central (or smallest\note{%
  We usually create a 68-percent (or 95-percent) interval or region by
  excluding the top and bottom 16 percent (or 2.5 percent) of the
  posterior samples.  Some more aggressive investigators find the
  smallest interval that contains 68 (or 95) percent of the samples.
  Both are legitimate, and they are near-identical for distributions
  that are near-symmetric around the mean.})
  interval of the parameter that contains 68 percent of the posterior
  samples.
The ``two-sigma'' error bar would be the same but for the 95-percent interval.
Again, these limits can be estimated%
  \note{One amusing thing about samplings,
  which are so beloved of us probabilistic (Bayesian) reasoners,
  is that anything we \emph{do} with a sampling,
  like estimate an integral or a quantile,
  is just an old-school frequentist estimator.
  A frequentist estimator of a Bayesian quantity, to be sure, but a frequentist estimator nonetheless.}
  by ordering the $q(\pars_k)$ values,
  and finding the $\pars$ values such that some fraction (0.68 or 0.95)
  of the samples lie between them.
Again, to form these limits, $q(\pars)$ must be a scalar function (so sorting is well defined).
Because they are so afraid of being confused with frequentists,
Bayesians often call these regions ``credible'' rather than
``confidence''\note{We don't like this ``credible'' terminology,
  though we sometimes adopt it to avoid confusion.}

>>>>

Given this pdf $p(\pars)$, we can define \emph{expectation values}
$E_{p(\pars)}[\pars]$
for $\pars$ or for any quantity that can be expressed as a function $g(\pars)$
of $\pars$:
\begin{eqnarray}
E_{p(\pars)}[\pars] &\equiv& \int \pars\,p(\pars)\,\dd \pars
\\
E_{p(\pars)}[g(\pars)] &\equiv& \int g(\pars)\,p(\pars)\,\dd \pars
\quad ,
\end{eqnarray}
where again the integrals are implicitly definite integrals over the
entire domain
of $\pars$ (all the parts of $\pars$ space in which $p(\pars)$ is finite) and the
integrals are multi-dimensional if $\pars$ is multi-dimensional.
These expectation values are the mean values of $\pars$ and $g(\pars)$ under
the pdf.  A good sampling---and really this is the definition of a good
sampling---makes the sampling approximation to these integrals
accurate.
With a good sampling $\setof{\pars_k}_{k=1}^K$ the integrals get replaced
with sums over samples $\pars_k$:
\begin{eqnarray}
E_{p(\pars)}[\pars] &\approx& \frac{1}{K}\,\sum_{k=1}^K \pars_k
\\
E_{p(\pars)}[g(\pars)] &\approx& \frac{1}{K}\,\sum_{k=1}^K g(\pars_k)
\quad .
\end{eqnarray}
That is, a sampling is good when any expectation value of interest is
accurately computed via the sampling approximation.
The word ``accurately'' here translates into some kinds of theorems
about limiting behavior; the general idea is that the sampling
approximation becomes exact as $K$ goes to infinity.
The size $K$ of the sampling you need \emph{in practice} will depend
on the expectations you want to compute, and the accuracies you need.

@@

When the sampling $\setof{\pars_k}_{k=1}^K$ is of one of these badly
normalized functions $f(\pars)$---as it usually will be---the
sampling-approximation expectation values are the expectation values
under the properly-normalized corresponding pdf, even though you might
\emph{never learn that normalization}.
When we run \MCMC\ sampling on $f(\pars)$, a function that differs from a
pdf $p(\pars)$ by some unknown normalization constant, then the sampling
permits computation of the following kinds of quantities:
\begin{eqnarray}
E_{p(\pars)}[g(\pars)] &\equiv& \frac{\int g(\pars)\,f(\pars)\,\dd \pars}{\int f(\pars)\,\dd \pars}
\\
E_{p(\pars)}[g(\pars)] &\approx& \frac{1}{K}\,\sum_{k=1}^K g(\pars_k)
\label{eq:expectationvalue}
\quad .
\end{eqnarray}
That is, the sampling can be constructed (as we will show below in \sectionname~\ref{sec:MH}) from
evaluations of $f(\pars)$ directly, and it permits you to compute expectation values
without ever requiring you to integrate either the numerator integral
or the denominator integral, both of which are generally
intractable.\note{As we will
  see, the intractability comes from various directions, but one is
  that the dimension of $\pars$ gets large in most realistic situations.
  Another is that the support of $p(\pars)$ tends to be, in normal
  inference situations, \emph{far smaller} than the full domain of
  $\pars$.
  That is, the pdf is at or very close to zero over all but some tiny
  and very hard-to-find part of the space.}

@@

That is, if the expectation value $E[g]$ is computed by the sampling
approximation given in \equationname~(\ref{eq:expectationvalue}),
there is an uncertainty on that expectation given by
\begin{eqnarray}
  \Var[E_{p(\pars)}[g(\pars)]] &=& \frac{\tau_q}{K}\,\Var_{p(\pars)}[g(\pars)]
  \\
  &\approx& \frac{\tau_q}{K^2}\,\sum_{k=1}^K \left(g(\pars_k) -
    E_{p(\pars)}[g(\pars)]\right)^2
  \quad,
\label{eq:tauvar}
\end{eqnarray}
where we are being a bit loose with terminology because this is the
variance not on $E_{p(\pars)}[g(\pars)]$ but on our sampling-based
\emph{estimator} for
$E_{p(\pars)}[g(\pars)]$, and the $E_{p(\pars)}[g(\pars)]$ inside the sum is
that estimator too.
It is slightly off-topic here, but related to this, if you have a long
chain of length $K$ and with autocorrelation time $\tau$ and you thin
the chain down to every $\tau$th sample (instead of using all $K$
samples), you will get just as good sampling-based estimates as if you
had used all $K$ samples.

It is also worth noting that there is not just one autocorrelation function or
autocorrelation time for an \MCMC\ chain.
Instead, there is an autocorrelation time $\tau_g$ for every function
$g(\pars)$ and these will, in general, be different.
Therefore, the autocorrelation time used in \equationname~(\ref{eq:tauvar}) is
the time computed on the chain of $g(\pars)$ values instead of the
autocorrelation time for $\pars$.

@@

For example,
  imagine that by \MCMC\ you have generated
  $K$ samples $\pars_k$ of a parameter vector $\pars$.
Now you have a scalar function $g(\pars)$ which takes the parameter vector
  and returns a scalar value%
  \note{For the median-of-sampling value, $g(\pars)$ needs to return a scalar,
  but technically, for the mean-of-sampling, $g(\pars)$ can be a vector or something high-dimensional.}
  which could be as simple as a single component of the parameter vector
  (one parameter from the list),
  or something more complex.
The mean-of-sampling value $\mean{q}$ is just
\begin{eqnarray}
\mean{q} &\leftarrow& \frac{1}{K}\,\sum_{k=1}^K g(\pars_k)
\quad ,
\end{eqnarray}
  and the median-of-sampling value is just the $[K/2]$th value of $g(\pars_k)$
  when the $g(\pars_k)$ have been ordered (sorted).
These are both produced by integrals;
  the first is the value returned by the expectation integral estimate,
  the second is the value past which half of the integrated pdf lies.

What are your best options for the ``uncertainty'' or ``limits'' on this parameter?
The only simple integral-based options are either variances or else posterior quantiles.
We usually use quantiles.
That is, the ``one-sigma'' error bar can be taken to be the half-size of the central (or smallest\note{%
  We usually create a 68-percent (or 95-percent) interval or region by
  excluding the top and bottom 16 percent (or 2.5 percent) of the
  posterior samples.  Some more aggressive investigators find the
  smallest interval that contains 68 (or 95) percent of the samples.
  Both are legitimate, and they are near-identical for distributions
  that are near-symmetric around the mean.})
  interval of the parameter that contains 68 percent of the posterior
  samples.
The ``two-sigma'' error bar would be the same but for the 95-percent interval.
Again, these limits can be estimated%
  \note{One amusing thing about samplings,
  which are so beloved of us probabilistic (Bayesian) reasoners,
  is that anything we \emph{do} with a sampling,
  like estimate an integral or a quantile,
  is just an old-school frequentist estimator.
  A frequentist estimator of a Bayesian quantity, to be sure, but a frequentist estimator nonetheless.}
  by ordering the $g(\pars_k)$ values,
  and finding the $\pars$ values such that some fraction (0.68 or 0.95)
  of the samples lie between them.
Again, to form these limits, $g(\pars)$ must be a scalar function (so sorting is well defined).
Because they are so afraid of being confused with frequentists,
Bayesians often call these regions ``credible'' rather than
``confidence''\note{We don't like this ``credible'' terminology,
  though we sometimes adopt it to avoid confusion.}
